RBE/CS549 Project: Boat Tracking/Detection

### 2017-12-14 paper/main.tex
Fix more 80-character lines
Write a bit on the report

The boat that the data is collected on is relatively stable
and the GoPro is mounted in the same spot for the length
of the video. As such, we should be able to safely ignore
the rotational pose of the boat (when we apply these
algorithms to our final boat, we will have to account for
the much more noticeable movement of that boat).

\begin{figure}[H]
\includegraphics[width=12cm]{sample1}
\centering
\caption{Sample image 1}
\end{figure}
\begin{figure}[H]
\includegraphics[width=12cm]{sample2}
\centering
\caption{Sample image 1}
\end{figure}

@@

\subsection{Naive K-Means Classifier}
We initially started with a basic two-class k-means classifier in order to
observe how well it would perform.  The classifier was trained on the first
image in the video sequence, and the l2 distance to the two cluster centers was
used to classify each pixel in subsequent frames.  Surprisingly enough, the
2-class k-means classifier worked relatively robustly on the videos that we
provided the algorithm, as show in figure \ref{fig:kmeans}.  The water pixels
are mostly labeled as one class, and pixels that not water are mostly labeled as
another class with a little bit of misclassification noise.  The boat and it's
wake are clearly labeled as potential obstacles to track, which is what we want.
The land areas are labeled as water pixels, but for our implementation we want
to ignore tracking the land anyways. There is a slight bit of noise in the water
reflections towards the right side of figure \ref{fig:kmeans}, but this naive
method still provides very useful information.

Kmeans does not always associate the same class identifier with the water and
non-water pixels each time, so we must perform an intermediary step to detect
which class identifier it associates with water pixels.  To do so, we find the
number of pixels in each class in the bottom half of the image, and use the sums
to vote for which class is associated with water pixels. This assumes that the
boat is upright, and that a majority of the pixels in the bottom half of the
image are water, and holds true for all of the cases we observed.  If the water
class ID is not what was expected, the cluster centers are swapped so that the
water class ID is always a known value.

\begin{figure}
\label{fig:kmeans}
\includegraphics[width=7.8cm]{hsv_kmeans2_orig}
\includegraphics[width=8.5cm]{hsv_kmeans2_result}
\centering
\caption{Naive K-Means Classifier Results}
\end{figure}

@@

\subsection{Sky Suppression}
The k-means classifier provides information about which pixels are water and
which pixels are not water, but we would also like the algorithm to only focus
on potential obstacles in the water.  The region of the sky shown in figure
\ref{fig:kmeans} is classified as a non-water pixel, and we would like to
suppress this response to avoid detecting the sky as an obstacle. With only the
wave reflection suppression confidences, most of the sky is suppressed, but
tends to have higher confidences near land as shown in figure
\ref{fig:skysuppress}.  Since the sky pixels are all connected, and define a
very large region, we used the Matlab bwareaopen function that removes binary
pixels in connected areas that are smaller than a particular size.  By looking
only at pixels that are classified as "not water", we can detect the sky region
by filtering for pixel clusters that are larger than 1/4 of the area of the
image, and set the confidence values for those pixels to zero to suppress them.
In some cases, potential objects to track were touching the sky region and the
bwareaopen results tended to remove those object to track.  To fix this, an
erode operation is done on the non-water class pixels to better separate the sky
and potential objects to track.

\begin{figure}
\label{fig:skysuppress}
\includegraphics[height=5cm]{hsv_confidence}
\includegraphics[height=5cm]{hsv_kmeans2_suppressed}
\centering
\caption{Sky Suppression Before and After}
\end{figure}

\subsection{Confidence Image Results}
When the above confidence values and suppressions described above are combined,
a final confidence value image is produced. Pixels that are more likely to be
objects to track have higher confidence values that pixels that are more likely
to be water areas. Figure \ref{fig:boatconf} shows the detector correctly
detecting a boat in the image.  Figure \ref{fig:towerconf} shows the detector
correctly detecting a tower in the distance.  Figure \ref{fig:buoyconf} shows
the detector identifying a buoy off in the distance.

\begin{figure}[H]
\label{fig:boatconf}
\includegraphics[width=7.8cm]{hsv_kmeans2_orig}
\includegraphics[width=8.5cm]{hsv_kmeans2_suppressed}
\centering
\caption{Boat Confidence Results}
\end{figure}

\begin{figure}[H]
\label{fig:towerconf}
\includegraphics[width=7.9cm]{hsv_kmeans2_tower}
\includegraphics[width=8.3cm]{hsv_kmeans2_tower_result}
\centering
\caption{Tower Confidence Results}
\end{figure}

\begin{figure}[H]
\label{fig:buoyconf}
\includegraphics[width=7.8cm]{hsv_kmeans2_buoy}
\includegraphics[width=8.5cm]{hsv_kmeans2_buoy_result}
\centering
\caption{Buoy Confidence Results}
\end{figure}

@@

\subsection{Expected Flow}

@@

\section{Blob Detection and Filtering}

\section{Object Matching}

@@

@@

>>>>

The boat that the data is collected on is relatively stable
and the GoPro is mounted in the same spot for the length
of the video. As such, we should be able to safely ignore
the rotational pose of the boat (when we apply these
algorithms to our final boat, we will have to account for
the much more noticeable movement of that boat).
\begin{figure}[H]
\includegraphics[width=12cm]{sample2}
\centering
\caption{Sample image 1}
\end{figure}

@@

\subsection{Naive K-Means Classifier}
We initially started with a basic two-class k-means classifier in order to
observe how well it would perform.  The classifier was trained on the first
image in the video sequence, and the L2 distance to the two cluster centers was
used to classify each pixel in subsequent frames.  Surprisingly enough, the
2-class k-means classifier worked relatively robustly on the videos that we
provided the algorithm, as show in figure \ref{fig:kmeans}.  The water pixels
are mostly labeled as one class, and pixels that not water are mostly labeled as
another class with a little bit of misclassification noise.  The boat and it's
wake are clearly labeled as potential obstacles to track, which is what we want.
The land areas are labeled as water pixels, but for our implementation we want
to ignore tracking the land anyways. There is a slight bit of noise in the water
reflections towards the right side of figure \ref{fig:kmeans}, but this naive
method still provides very useful information.

K-means does not always associate the same class identifier with the water and
non-water pixels each time, so we must perform an intermediary step to detect
which class identifier it associates with water pixels.  To do so, we find the
number of pixels in each class in the bottom half of the image, and use the sums
to vote for which class is associated with water pixels. This assumes that the
boat is upright, and that a majority of the pixels in the bottom half of the
image are water, and holds true for all of the cases we observed.  If the water
class ID is not what was expected, the cluster centers are swapped so that the
water class ID is always a known value.

\begin{figure}
\includegraphics[width=7.8cm]{hsv_kmeans2_orig}
\includegraphics[width=8.5cm]{hsv_kmeans2_result}
\centering
\caption{Naive K-Means Classifier Results}
\label{fig:kmeans}
\end{figure}

@@

\subsection{Sky Suppression}
The k-means classifier provides information about which pixels are water and
which pixels are not water, but we would also like the algorithm to only focus
on potential obstacles in the water.  The region of the sky shown in figure
\ref{fig:kmeans} is classified as a non-water pixel, and we would like to
suppress this response to avoid detecting the sky as an obstacle. With only the
wave reflection suppression confidences, most of the sky is suppressed, but
tends to have higher confidences near land as shown in figure
\ref{fig:skysuppress}.  Since the sky pixels are all connected, and define a
very large region, we used the Matlab \texttt{bwareaopen} function that removes binary
pixels in connected areas that are smaller than a particular size.  By looking
only at pixels that are classified as "not water", we can detect the sky region
by filtering for pixel clusters that are larger than 1/4 of the area of the
image, and set the confidence values for those pixels to zero to suppress them.
In some cases, potential objects to track were touching the sky region and the
\texttt{bwareaopen} results tended to remove those object to track.  To fix this, an
erode operation is done on the non-water class pixels to better separate the sky
and potential objects to track.

\begin{figure}
\includegraphics[height=5cm]{hsv_confidence}
\includegraphics[height=5cm]{hsv_kmeans2_suppressed}
\centering
\caption{Sky Suppression Before and After}
\label{fig:skysuppress}
\end{figure}

\subsection{Confidence Image Results}
When the above confidence values and suppressions described above are combined,
a final confidence value image is produced. Pixels that are more likely to be
objects to track have higher confidence values than pixels that are more likely
to be water areas. Figure \ref{fig:boatconf} shows the detector correctly
detecting a boat in the image.  Figure \ref{fig:towerconf} shows the detector
correctly detecting a tower in the distance.  Figure \ref{fig:buoyconf} shows
the detector identifying a buoy off in the distance.

\begin{figure}[H]
\includegraphics[width=7.8cm]{hsv_kmeans2_orig}
\includegraphics[width=8.5cm]{hsv_kmeans2_suppressed}
\centering
\caption{Boat Confidence Results}
\label{fig:boatconf}
\end{figure}

\begin{figure}[H]
\includegraphics[width=7.9cm]{hsv_kmeans2_tower}
\includegraphics[width=8.3cm]{hsv_kmeans2_tower_result}
\centering
\caption{Tower Confidence Results}
\label{fig:towerconf}
\end{figure}

\begin{figure}[H]
\includegraphics[width=7.8cm]{hsv_kmeans2_buoy}
\includegraphics[width=8.5cm]{hsv_kmeans2_buoy_result}
\centering
\caption{Buoy Confidence Results}
\label{fig:buoyconf}
\end{figure}

@@

\subsection{Expected Flow} \label{expected_flow}

@@

\section{Object Creation and Management}

\subsection{Blob Detection and Filtering}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{threshold_bwlabel}
\caption{Result of thresholding after dilation/labelling}
\label{fig:threshold_bwlabel}
\end{figure}

From the results in Figures \ref{fig:boatconf} and
\ref{fig:diff_magnitude}, we have a set of confidence values that can be used to
identify where individual obstacles may be. In order to combine the two, we
simply scale each result by some constant and sum the two to provide a
detection score. Given this score, we then:
\begin{enumerate}
\item Threshold against some constant (tunable) value
\item Do an erode operation using a circular erosion pattern
\item Do a dilate in reverse, using a radius roughly twice that of the erosion,
      so that not only is lost space taken up but nearby segments are connected.
\item Label the connected regions using \texttt{bwlabel} (corresponding with
      Figure \ref{fig:threshold_bwlabel}).
\item For each of these regions, throw out any regions that meet any of the
      following conditions (resulting in Figure \ref{fig:filtered_labels}:
  \begin{itemize}
    \item The object is too ``blue'' (by comparing to some hue-based value,
          in order to try and filter out any water that got detected)
    \item The object is too small (suggesting it may not really exist)
    \item The object is too large (sometimes large (~half the image) connected
          segments will appear that are virtually never real objects)
    \item It is moving too slowly relative to us (relative to the camera,
          the only things that should be that constant are on our
          boat---anything else should at least contain some noise)
  \end{itemize}
\end{enumerate}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{filtered_labels}
\caption{Result of filtering the blobs in Figure \ref{fig:threshold_bwlabel},
         with the boat and parts of its wake remaining}
\label{fig:filtered_labels}
\end{figure}

Following these steps, we initialize each remaining blob as an object, reducing
it to a bounding box (rather than an exact set of pixels), a median velocity
(from the optical flow), a position (the center of the bounding box), a label
(arbitrarily assigned), and a confidence score (initialized to 0.2, to represent
an object that we are not entirely sure exists yet).

We then go through and, for each object, attempt to both associate it with other
newly created objects (in case any objects were over-segmented) as well as
objects that we've kept track of from the previous frame.

\subsection{Object Matching}

When attempting to match objects, we first update all the old objects so that
their positions have been shifted by their velocities, to better match objects
that are moving between frames. Furthermore, we reduce the bounding box sizes on
old objects to reflect a greater uncertainty; if the object really does exist
over the entire previous bounding box, then that should be reaffirmed by any
blobs detected in the current frame.

We then perform the following steps:

\begin{enumerate}
\item For each pair of new objects, as well as each pairing of one new and one
      old object, we compute a difference score that is the sum of:
  \begin{itemize}
  \item The difference in velocity direction
  \item One minus the quotient in velocity magnitudes (speed), such that the
        larger of the two speeds shall always be in the denominator
  \item The difference in the positions (L2-norm) of the two objects, scaled
        along each axis by the mean width and mean height between the two
        objects (such that if the objects are perfectly aligned in one dimension
        and touching edges in the other, they will be a distance of 1.0 apart).
  \end{itemize}
\item We threshold every difference, leaving a matrix of object ``neighbors''.
\item Iterating over the newly created objects (in arbitrary order), we
      combine that object with it's immediate neighbors, and then remove
      all the objects that were combined from future
      consideration\footnote{Ideally, we might do some sort of search where we
      combine neighbors-of-neighbors and so on, but for simplicities' sake, and
      to avoid situations where we end up combining large numbers of objects
      that all are tenuously connected, we just explore one step out from any
      given object.}.
\item In order to combine each set of objects that we associated, we:
  \begin{itemize}
    \item Create a bounding box that surrounds all of the combined objects
    \item Give the amalgamation the label (and time-of-existence) of the oldest
          object being included, on the assumption that we are interested
          in keeping tracking going across the longest period of time possible.
    \item Perform a weighted average of the velocities, weighting by object area.
    \item Take the maximum of the confidence scores for the individual objects,
          assign it to the result object, and then increment it up slightly on
          the assumption that any time we are combining more than one object,
          we are probably becoming more certain in the overall object's
          existence.
  \end{itemize}
\item Any old objects that were not associated with blobs in the current frame
      have their confidence scores reduced and, if they dip below a confidence
      of 0, are deleted permanently.
\item Finally, for every object being tracked, we keep a history of its last 20
      positions. Once an object accumulates a history of more than around 5
      frames (to reduce potential for individual frames causing too much noise),
      we start to compare the object's overall average position change to its
      velocity. If the two differ by more than a particular threshold, then we
      reduce the confidence score again. This helps to mitigate object resulting
      from the wake of the boat, as such objects generally have optical flows
      inconsistent with where the object actually moves frame-to-frame.
\end{enumerate}

\section{Objective Evaluation}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{example_detection}
\caption{Sample frame of the whole stack running}
\label{fig:example_detection}
\end{figure}

In Figure \ref{fig:example_detection}, we see an individual frame from the
algorithm, where we have:
\begin{itemize}
\item The boat identified by a yellow (indicating that it has existed for at
      least 10 frames) solid (indicating high confidence) box with an arrow
      pointing leftwards, roughly in the direction of its movement.
\item Several solid and dashed boxes detecting the wake of the boat.
\end{itemize}

Longer videos can be found in our presentation:

\begin{description}
\item[\url{https://drive.google.com/file/d/1FoKWkp5xDSKwp5YyluKrclo_kGFU4gsg/view?usp=sharing}]
  In this video we have a longer view of the same boat as Figure
  \ref{fig:example_detection}
\item[\url{https://drive.google.com/file/d/1ndVSwF77_4va6U8rWT4FGJOXq1AjKL_Z/view?usp=sharing}]
  This has a slightly shorter capture of a different boat, indicating some
  degree of robustness to different individual instances
  \ref{fig:example_detection}
\end{description}

\subsection{Timing}

As we are interested in ultimately implementing some of these algorithms on our
actual robot, we are interested in the run time of our algorithms. These
are extremely rough numbers, from running only single-threaded on a laptop CPU,
but were reasonably consistent across runs:
\begin{itemize}
\item Around 3 seconds total per frame
\item Around 1 second to call \texttt{opticalFlowFarneback} on the full
      1920x1080 image
\item Around half a second to perform the linear fit described in section
      \ref{expected_flow}
\item Around a second and a half spread out among other steps
\end{itemize}

The linear fit step in particular initially took up a substantial amount of
time, but we discovered that simply subsampling the data and only using around
10 percent of it worked almost as well while reducing the runtime to the half
second seen here.

The optical flow timing is exacerbated by the fact that we are running optical
flow on the entire, reasonably high-resolution, image. Were we to focus better
on interest regions, or simply use a lower resolution, we could likely achieve
substantially better results.

\subsection{Robustness}

\begin{figure}[H]
\includegraphics[width=12cm]{sample1}
\centering
\caption{Mirror-like water}
\label{fig:mirror}
\end{figure}

While the algorithms do work reasonably well for footage from these particular
conditions, we had troubles processing some parts of the data. In particular,
early in certain mornings when not only was the sun interfering with the camera,
but also in situations such as Figure \ref{fig:mirror}, in which the water is
relatively flat (reducing texture) and highly reflective, meaning that it is no
longer particularly blue.

@@

For future work, we need to actually improve the timing (as mentioned above) to
work in real time, improve robustness for a greater variety of conditions, and
to increase the generality to function on a less stable platform--our boat will
certainly be pitching and rolling substantially more than the platform we had
for this footage.
